Detecting Hate Speech with GPT-3

### 2022-03-16 outputs/documents/paper.tex
Final read of paper

\documentclass[12pt,]{article}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

@@

\setcounter{secnumdepth}{0}

@@

\title{Detecting Hate Speech with GPT-3 \thanks{Code and data are available at: \url{https://github.com/kelichiu/GPT3-hate-speech-detection}. We gratefully acknowledge the support of Gillian Hadfield and the Schwartz Reisman Institute for Technology and Society. We thank Amy Farrow, Haoluan Chen, Mauricio Vargas Sepúlveda, and Tom Davidson for helpful suggestions. Comments on the 22 March 2021 version of this paper are welcome at: \href{mailto:rohan.alexander@utoronto.ca}{\nolinkurl{rohan.alexander@utoronto.ca}}.}  }

\author{\Large Ke-Li Chiu\vspace{0.05in} \newline\normalsize\emph{University of Toronto}   \and \Large Rohan Alexander\vspace{0.05in} \newline\normalsize\emph{University of Toronto and Schwartz Reisman Institute}  }

@@

@@

% move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Ke-Li Chiu (University of Toronto) and Rohan Alexander (University of Toronto and Schwartz Reisman Institute)},
             pdfkeywords = {GPT-3; natural language processing; quantitative analysis; hate speech.},  
            pdftitle={Detecting Hate Speech with GPT-3},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

@@

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

@@

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Ke-Li Chiu} \hskip 15pt \emph{\small University of Toronto}   \par \textbf{\authorfont Rohan Alexander} \hskip 15pt \emph{\small University of Toronto and Schwartz Reisman Institute}

@@

\noindent Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist? We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 is able to identify sexist or racist text with an accuracy between 48 per cent and 69 per cent. With few-shot learning and an instruction included in the prompt, the model's accuracy can be as high as 78 per cent. We conclude that large language models have a role to play in hate speech detection, and that with further development language models could be used to counter hate speech and even self-police.

@@

\vskip -8.5pt

 % removetitleabstract

\noindent

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Sophisticated natural language processing (NLP) models, such as OpenAI's GPT-3, can produce hateful text. In particular, there have been many examples of text being generated that targets marginalized groups based on their sex, race, sexual orientation, and other characteristics. Large language models are trained on enormous datasets from various sources. This means that untruthful statements, human biases, and abusive language are inevitably included. Further, \citet{bender2021dangers} describes how even when the models do not possess intent, they risk producing synthetic texts that are offensive or discriminatory and thus cause unpleasant, or even triggering, interaction experiences.

The sources of the training datasets raise concerns around three issues: exclusion, over-generalization, and exposure \citep{hovy2016social}. Exclusion happens due to the demographic bias in the dataset. In the case of language models that are trained on US and UK English scraped from the Internet, datasets may be disproportionately white, male, and young. Therefore, it is not surprising to see white supremacist, misogynistic, and ageist content being over-represented in training datasets \citep{bender2021dangers}. Over-generalization stems from the assumption that what we see in the dataset represents what occurs in the real world. Words such as `always', `never', `everybody', or `nobody' are frequently used for rhetorical purpose instead of their literal meanings. However, language models do not always recognize this and make inference based on generalized statements using these words. For instance, hate speech commonly uses generalized language for targeting a group such as `all' and `every', and a model trained on these statements may generate similarly generalized and harmful statements. Finally, exposure refers to the relative attention, and hence considerations of importance, given to something. In the context of NLP this may be reflected in the emphasis on English-language created under particular circumstances, rather than another languages or circumstances that may be more prevalent.

While these, and other, issues give us pause, the dual-use problem, which is that the same technology can be applied to both good and bad uses, provides motivation. For instance, while stylometric analysis can reveal the identity of political dissenters, it can also solve the unknown authorship of historic text \citep{hovy2016social}. In this paper we are interested in whether given large language models can produce harmful language, can they also identify, or learn to identify, harmful language?

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of harmful languages can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. In particular we consider 120 different extracts that have been categorized as `racist', `sexist', or `neither' in the single-category settings and 240 different extracts in the mixed-category settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with few-shot learning when an instruction is included. In that setting the model is able to accurately classify around 78 per cent of the extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

@@

In its simplest form, a language model is a probability distribution over sequences of words, which are usually generalized further to tokens. The sequence of tokens constitutes different linguistic units --- words, sentences, and even documents \citep{bengio2003neural}. Language models predict the next token based on inputs. If we consider each token in a vocabulary as a dimension, then the dimensionality of language quickly becomes large \citep{rosenfeld2000two}. Over time a variety of statistical language models have been created to nonetheless enable prediction. The n-gram is one of the earliest language models. It works by considering the co-occurrence of tokens in a sequence. In the early 2000s, prominent neural network language models were developed, for instance \citet{bengio2003neural}. These were then built on by word embeddings language models in the 2010s in which the distance between tokens represents how related those tokens are, for instance \citet{turian2010word}. In 2017, \citet{vaswani2017attention} introduced the Transformer, which marked a new era for language models. The Transformer is a network architecture for neural networks that can be trained more quickly than many other approaches \citep{vaswani2017attention}. Now most of the representative pre-trained language models, such as BERT \citep{devlin2018bert}, GPT-2 \citep{radford2019language}, and GPT-3 \citep{brown2020language}, are built on this architecture.

GPT-3 is the third generation of the Generative Pre-trained Transformer models created by OpenAI. Until January 11, when \citet{fedus2021switch} announced a Transformer model with a trillion parameters, GPT-3 was the largest, publicly-known, Transformer language model. GPT-3 is distinctive from its predecessors because of few-shot learning. This means that GPT-3 can `learn' to perform a new task based on only a few examples, expressed in natural language, instead of a fine-tuning process that can require a large amount of data. GPT-3 has led to unexpected NLP applications, such as computational code generation given natural language prompts. However, like other language models, GPT-3 has also generated inappropriate or even hateful content. For instance, \citet{mcguffie2020radicalization} demonstrated the use of GPT-3 in mass-producing radicalized text targeting the Islamic population.

@@

Hate speech detection is of interest to researchers in a variety of domains including computer science and sociology, as well as industry and the judiciary. Detecting hate speech is difficult because the definition of hate speech varies depending on the complex intersection of the topic of the assertion, the context, the timing of the post, synchronized world events, and the identity of speaker and recipient \citep{schmidt2017survey}. Moreover, it is difficult to discern hate speech from merely offensive language \citep{davidson2017automated}. Since hate speech is prohibited in several countries, misclassification of hate speech can become a legal problem. For instance, in Canada, speech that contains `public incitement of hatred' or `wilful promotion of hatred' is specified by the Criminal Code \citep{act2021justice}. Policies toward hate speech are more detailed in some social media platforms. For instance, the Twitter Hateful Conduct Policy states:

@@

There has been a large amount of research focused on detecting hate speech. And as part of this various hate speech datasets have been created and examined. For instance, \citet{waseem2016hateful} create a dataset that captures hate speech in the form of racist and sexist language that includes domain expert annotation, and \citet{davidson2017automated} trains a classifier to distinguish between hate speech and offensive language. And it is important to note that even these datasets have bias. For instance, \citet{davidson2019racial} found racial bias in five different sets of Twitter data annotated for hate speech and abusive language. They found that tweets written in African American English are more likely to be labeled as abusive.

@@

We examine the ability of GPT-3 to identity hate speech in zero-shot, one-shot, and few-shot settings. There are a variety of parameters, such as temperature, that control the degree of text variation. To enhance consistency, the temperature is set to zero in our experiments. There are two categories of hate speech that are of interest in this paper. The first targets the race of the recipient, and the other targets the gender of the recipient. With zero- and one-shot learning, the model identifies hate speech one category at a time. With few-shot learning, the categories are mixed, and the model is asked to classify an input as sexist, racist, or neither.

@@

We use the ETHOS dataset created by \citet{mollas2020ethos}. ETHOS is based on comments found in YouTube and Reddit. The ETHOS YouTube data is collected through Hatebusters \citep{anagnostou2018hatebusters}. Hatebusters is a platform that collects comments from YouTube and assigns a `hate' score to them using a support vector machine. The Reddit data is collected from the Public Reddit Data Repository \citep{baumgartner2020pushshift}. The dataset has two variants: binary and multi-label. In the binary dataset comments are classified as hate or non-hate. In the multi-label variant, the comments are evaluated on measures that include violence, gender, race, ability, religion, and sexual orientation. The examples provided in this paper are from the ETHOS dataset and hence contain typos, misspelling, and offensive content.

There are 998 statements in the ETHOS dataset that have a binary classification of hate speech or not hate speech. Of these, the 433 statements that contain hate speech additionally have multiple labels that classify the content, for instance, does the comment have to do with violence, gender, race, nationality, disability, etc. We considered the 136 statements that contain race-based hate speech, and we focus on the 76 whose race-based score is at least 0.5. Similarly, we considered the 174 statements that contain gender-based hate speech, and again focused on the 84 whose gender-based score is at least 0.5. We randomly draw 120 of these statements, using stratified sampling, to create a balanced dataset of 60 sexist statements and 60 racist statements. For comparison purposes, we randomly draw 120 statements that are classified as not hate speech.

@@

@@

Under the few-shot setting, three examples are provided to the model. There are two settings considered in the few-shot experiments: single-category, and mixed-category. In the single-category setting, the model is given two examples that belong to the selected category, and one example that is not hate speech. The answer is expected to be binary. For instance, the model is expected to categorize a passage of text as `sexist' or `not-sexist'. An example of this is:

@@

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as `sexist', `racist', or `neither'. There are ten sets of examples; each set of examples is used for the model to classify 60 racist comments, 60 sexist comments, and 120 comments that are in neither category This results in 2,400 responses in total. The ten sets of examples are available in Table \ref{tab:appendixa} in the Appendix. An example of this is:

@@

\hypertarget{result}{%
\section{Result}\label{result}}

@@

The results of the zero-shot experiments are presented in Table \ref{tab:zeroshot}. The model has 35 matches and 25 mismatches in the sexist category, and 23 matches and 37 mismatches in the racist category. The model performs better when identifying sexist comments compared with identifying racist comments. However, the overall ratio of matches and mismatches is 58:62. In other words, the accuracy in identifying hate speech in the zero-shot setting is 48.3 per cent.

@@

\caption{\label{tab:zeroshot}Classification of statements with zero-shot learning}
\centering
\begin{tabular}[t]{llr}
\toprule
Result & Category & Count\\
\midrule
Match & Racist & 23\\
Match & Sexist & 35\\
Mismatch & Racist & 37\\
Mismatch & Sexist & 25\\
\bottomrule
\end{tabular}
\end{table}

@@

The results of the one-shot learning experiments are presented in Table \ref{tab:oneshot}. The model has 46 matches and 14 mismatches in the racist category, and 37 matches and 23 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model performs slightly better in identifying racist comments compared with identifying sexist comments. The general performance in the one-shot setting is also better than in the zero-shot setting. The overall ratio of matches and mismatches is 83:37. In other words, the accuracy of identifying hate speech in the one-shot setting is 69.2 per cent.

@@

\caption{\label{tab:oneshot}Classification of statements with one-shot learning}
\centering
\begin{tabular}[t]{llr}
\toprule
Result & Category & Count\\
\midrule
Match & Racist & 46\\
Match & Sexist & 37\\
Mismatch & Racist & 14\\
Mismatch & Sexist & 23\\
\bottomrule
\end{tabular}
\end{table}

@@

The results of the single-category, few-shot learning, experiments are presented in Table \ref{tab:fewshotsingle}. The model has 41 matches and 19 mismatches in the racist category, and 42 matches and 18 mismatches in the sexist category. Here, the model performs almost equally well at identifying racist comments compared with identifying sexist comments. The general performance in the few-shot learning setting is similar to the performance in the one-shot learning setting. The overall ratio of matches and mismatches is 83:37, or around 69.2 per cent, however the composition is different, compared with the one-shot learning setting.

@@

\caption{\label{tab:fewshotsingle}Classification of statements with single-category few-shot learning}
\centering
\begin{tabular}[t]{llr}
\toprule
Result & Category & Count\\
\midrule
Match & Racist & 41\\
Match & Sexist & 42\\
Mismatch & Racist & 19\\
Mismatch & Sexist & 18\\
\bottomrule
\end{tabular}
\end{table}

@@

To reduce the chance of the model generating answers that are out of scope, a brief instruction is added to the prompt, specifying that the answers be: `sexist', `racist', or `neither'. The addition of an instruction successfully restricts the generated answers within the specified terms. The unique generated answers are listed in Table \ref{tab:fewshotmixedanswerswithinstruct}.

>>>>

\documentclass[12pt,]{article}
\usepackage{amsmath}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

@@

% This controls whether you get numbers in the sections or not
% \setcounter{secnumdepth}{0}

@@

\title{Detecting Hate Speech with GPT-3 \thanks{Code and data are available at: \url{https://github.com/kelichiu/GPT3-hate-speech-detection}. We gratefully acknowledge the support of Gillian Hadfield, the Schwartz Reisman Institute for Technology and Society, and OpenAI for providing access to GPT-3 under the academic access program. We thank two anonymous reviews and the editor, as well as Amy Farrow, Haoluan Chen, John Giorgi, Mauricio Vargas Sepúlveda, Monica Alexander, Noam Kolt, and Tom Davidson for helpful discussions and suggestions. Please note that we have added asterisks to racial slurs and other offensive content in this paper, however the inputs and outputs did not have these. Comments on the 16 March 2022 version of this paper are welcome at: \href{mailto:rohan.alexander@utoronto.ca}{\nolinkurl{rohan.alexander@utoronto.ca}}.}  }

\author{\Large Ke-Li Chiu\vspace{0.05in} \newline\normalsize\emph{University of Toronto}   \and \Large Annie Collins\vspace{0.05in} \newline\normalsize\emph{University of Toronto}   \and \Large Rohan Alexander\vspace{0.05in} \newline\normalsize\emph{University of Toronto and Schwartz Reisman Institute}  }

@@

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Ke-Li Chiu (University of Toronto) and Annie Collins (University of Toronto) and Rohan Alexander (University of Toronto and Schwartz Reisman Institute)},
             pdfkeywords = {GPT-3; natural language processing; quantitative analysis; hate speech.},  
            pdftitle={Detecting Hate Speech with GPT-3},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

@@

@@

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Ke-Li Chiu} \hskip 15pt \emph{\small University of Toronto}   \par \textbf{\authorfont Annie Collins} \hskip 15pt \emph{\small University of Toronto}   \par \textbf{\authorfont Rohan Alexander} \hskip 15pt \emph{\small University of Toronto and Schwartz Reisman Institute}

@@

\noindent Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist. We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 can identify sexist or racist text with an accuracy between 57 per cent and 68 per cent depending on the category of text and type of learning. With few-shot learning, the model's accuracy can be as high as 88 per cent. Large language models have a role to play in hate speech detection, and with further development could eventually be used to counter hate speech.

@@

\vskip 6.5pt

\noindent  \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\textbf{This paper contains language and themes that are offensive.}

Natural language processing (NLP) models use normal words, often written text, as their data. For instance, one might have content from many books and want to group them into themes. Sophisticated NLP models are being increasingly embedded in society. For instance, Google Search uses an NLP model called BERT to better understand what is meant by a word given its context. Some sophisticated NLP models, such as OpenAI's Generative Pre-trained Transformer 3 (GPT-3), can additionally produce text as an output.

The text produced by sophisticated NLP models can be hateful. In particular, there have been many examples of text being generated that targets marginalized groups based on their sex, race, sexual orientation, and other characteristics. For instance, `Tay' was a Twitter chatbot released by Microsoft in 2016. Within hours of being released, some of its tweets were sexist. Large language models are trained on enormous datasets from various, but primarily internet-based, sources. This means they usually contain untruthful statements, human biases, and abusive language. Even though models do not possess intent, they do produce text that is offensive or discriminatory, and thus cause unpleasant, or even triggering, interactions \citep{bender2021dangers}.

Often the datasets that underpin these models consist of, essentially, the whole public internet. This source raises concerns around three issues: exclusion, over-generalization, and exposure \citep{hovy2016social}. Exclusion happens due to the demographic bias in the dataset. In the case of language models that are trained on US and UK English scraped from the Internet, datasets may be disproportionately white, male, and young. Therefore, it is not surprising to see white supremacist, misogynistic, and ageist content being over-represented in training datasets \citep{bender2021dangers}. Over-generalization stems from the assumption that what we see in the dataset represents what occurs in the real world. Words such as `always', `never', `everybody', or `nobody' are frequently used for rhetorical purpose instead of their literal meanings. But NLP models do not always recognize this and make inference based on generalized statements using these words. For instance, hate speech commonly uses generalized language for targeting a group such as `all' and `every', and a model trained on these statements may generate similarly generalized and harmful statements. Finally, exposure refers to the relative attention, and hence consideration of importance, given to something. In the context of NLP this may be reflected in the emphasis on English-language created under particular circumstances, rather than another language or circumstances that may be more prevalent.

While these issues, among others, give us pause, the dual-use problem, which is that the same technology can be applied for both good and bad uses, provides motivation. For instance, while stylometric analysis can reveal the identity of political dissenters, it can also solve the unknown authorship of historic text \citep{hovy2016social}. In this paper we are interested in whether large language models, given that they can produce harmful language, can also identify, or learn to identify, harmful language.

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of hateful language can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. We consider 120 different extracts that have been categorized as `racist', `sexist', or `neither' in single-category settings (zero-shot, one-shot, and few-shot) and 243 different extracts in mixed-category few-shot settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with mixed-category few-shot learning. In that setting the model can accurately classify around 86 per cent of the racist extracts and 88 per cent of sexist extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

The remainder of this paper is structured as follows: Section \ref{background} provides background information about language models and GPT-3 in particular. Section \ref{methods} introduces our dataset and our experimental approach to zero-, one-, and few-shot learning. Section \ref{results} conveys the main findings of those experiments. And Section \ref{discussion} adds context and discusses some implications, next steps, and weaknesses. Appendices \ref{appendxa}, \ref{appendxb}, and \ref{appendxc} contain additional information.

@@

In its simplest form, a language model involves assigning a probability to a certain sequence of words. For instance, the sequence `the cat in the hat' is probably more likely than `the cat in the computer'. We typically talk of tokens, or collections of characters, rather than words, and a sequence of tokens constitutes different linguistic units: words, sentences, and even documents \citep{bengio2003neural}. Language models predict the next token based on inputs. If we consider each token in a vocabulary as a dimension, then the dimensionality of language quickly becomes large \citep{rosenfeld2000two}. Over time a variety of statistical language models have been created to nonetheless enable prediction. The n-gram is one of the earliest language models. It works by considering the co-occurrence of tokens in a sequence. For instance, given the four-word sequence, `the cat in the', it is more likely that the fifth word is `hat' rather than `computer'. In the early 2000s, language models based on neural networks were developed, for instance \citet{bengio2003neural}. These were then built on by word embeddings language models in the 2010s in which the distance between tokens represents how related those tokens are, for instance \citet{turian2010word}. In 2017, \citet{vaswani2017attention} introduced the Transformer, which marked a new era for language models. The Transformer is a network architecture for neural networks that can be trained more quickly than many other approaches \citep{vaswani2017attention}. Now most representative pre-trained language models, such as Google's Bidirectional Encoder Representations from Transformers (BERT) \citep{devlin2018bert}, as well as OpenAI's Generative Pre-trained Transformer (GPT)-2 \citep{radford2019language}, and GPT-3 \citep{brown2020language}, are built on this architecture. These models are widely used, for instance BERT is used by Google search.

GPT-3 is the third generation of the Generative Pre-trained Transformer models created by OpenAI, a private company in California that develops artificial intelligence models. GPT-3 is an autoregressive NLP model that can perform a variety of tasks, including responding to questions, summarizing, and parsing text, translation, and classification. Interactions with the model involve inputting some text as a prompt and GPT-3 returning a text completion according to that prompt.

GPT-3 is one of the largest, publicly available, Transformer language models. GPT-3 is distinctive to its predecessors because of few-shot learning. This means that GPT-3 can `learn' to perform a new task based on only a few examples, expressed in natural language, instead of a fine-tuning process that can require a large amount of data. GPT-3 has led to unexpected NLP applications, such as computational code generation given natural language prompts.

Like other language models, GPT-3 has also generated inappropriate or even hateful content. For instance, \citet{mcguffie2020radicalization} demonstrated the use of GPT-3 in mass-producing radicalized text targeting the Islamic population. And \citet{lin2021truthfulqa} show that GPT-3 and similar language models can propagate misconceptions that could deceive human readers. For instance, when asked `Who really caused 9/11?', they found that GPT-3 provided the false statement `The US government caused 9/11.'

@@

There is no commonly held definition of hate speech. Different legal jurisdictions have different definitions, as do different companies and other groups. One definition is `the intentional verbalization of prejudice against a social group' \citep{kennedy2018gab}. Detecting hate speech is difficult because the definition of hate speech varies depending on a complex intersection of the topic of the assertion, the context, the timing, outside events, and the identity of speaker and recipient \citep{schmidt2017survey}. Moreover, it is difficult to distinguish hate speech from offensive language \citep{davidson2017automated}. Hate speech detection is of interest to academic researchers in a variety of domains including computer science \citep{addressinghatespeech} and sociology \citep{davidson2017automated}. It is also of interest to industry, for instance to maintain standards on social networks, and in the judiciary to help identify and prosecute crimes. Since hate speech is prohibited in several countries, misclassification of hate speech can become a legal problem. For instance, in Canada, speech that contains `public incitement of hatred' or `wilful promotion of hatred' is specified by the Criminal Code \citep{act2021justice}. Policies toward hate speech are more detailed in some social media platforms. For instance, the Twitter Hateful Conduct Policy states:

@@

There has been a large amount of research focused on detecting hate speech. As part of this process, various hate speech datasets have been created and examined. For instance, \citet{waseem2016hateful} detail a dataset that captures hate speech in the form of racist and sexist language that includes domain expert annotation. They use Twitter data, and annotate 16,914 tweets: 3,383 as sexist, 1,972 as racist, and 11,559 as neither. They had a high degree of annotator agreement. Most of the disagreements were to do with sexism, and often explained by an annotator lacking apparent context. \citet{davidson2017automated} train a classifier to distinguish between hate speech and offensive language. To define hate speech, they use an online `hate speech lexicon containing words and phrases identified by internet users as hate speech'. Even these datasets have bias. For instance, \citet{davidson2019racial} found racial bias in five different sets of Twitter data annotated for hate speech and abusive language. They found that tweets written in African American English are more likely to be labeled as abusive.

@@

We examine the ability of GPT-3 to identify hate speech in zero-shot, one-shot, and few-shot settings. There are a variety of parameters, such as temperature, that control the degree of text variation. Temperature is a hyper-parameter between zero and one. Lower temperatures mean that the model places more weight on higher-probability tokens. To enhance consistency, the temperature is set to zero in our experiments and not varied. There are two categories of hate speech that are of interest in this paper. The first targets the race of the recipient, and the other targets the gender of the recipient. With zero-, one-, and few-shot single-category learning, the model identifies hate speech one category at a time. With few-shot mixed-category learning, the categories are mixed, and the model is asked to classify an input as sexist, racist, or neither. Zero-shot learning means an example is not provided in the prompt. One-shot learning means that one example is provided, and few-shot means that two or more examples are provided.

@@

We use the ETHOS dataset of \citet{mollas2020ethos}. ETHOS is based on comments from YouTube and Reddit. The ETHOS YouTube data is collected through Hatebusters \citep{anagnostou2018hatebusters}. Hatebusters is a platform that collects comments from YouTube and assigns a `hate' score to them using a support vector machine. That hate score is only used to decide whether to consider the comment further or not. The Reddit data is collected from the Public Reddit Data Repository \citep{baumgartner2020pushshift}. The classification is done by contributors to a crowd-sourcing platform. They are first asked whether an example contains hate speech, and then, if it does, whether it incites violence and other additional details. The dataset has two variants: binary and multi-label. In the binary dataset comments are classified as hate or non-hate based. In the multi-label variant comments are evaluated on measures that include violence, gender, race, ability, religion, and sexual orientation. The dataset that we use is as provided by the ETHOS dataset and hence contain typos, misspelling, and offensive content.

We begin with all of the 998 statements in the ETHOS dataset that have a binary classification of hate speech or not hate speech. Of these, the 433 statements that contain hate speech additionally have labels that classify the content. For instance, does the comment have to do with violence, gender, race, nationality, disability, etc. We initially considered all of the 136 statements that contain race-based hate speech, but we focus on the 76 whose race-based score is at least 0.5, meaning that at least 50 per cent of annotators agreed. Similarly, we initially considered all of the 174 statements that contain gender-based hate speech, and again focused on the 84 whose gender-based score is at least 0.5. To create a balanced dataset, we select 120 of these statements to result in 60 sexist statements and 60 racist statements. For comparison purposes, we randomly draw 120 statements that are classified as not hate speech. The balanced sample enables us to more easily understand our results.

For the purposes of one-shot and single-category few-shot learning, the dataset was split in half such that 30 racist comments, 30 sexist comments, and 60 neutral comments were used as inputs and the remainder were used as examples. This division was kept consistent for zero-shot learning to allow for the same set of 120 comments to be classified by all three approaches. With the introduction of the 10 static example sets for multi-category few-shot learning, 240 comments were classified for each example set. The comments in the example sets were drawn from the existing pool of 240 comments. For each example set, the comment from each category being used in the example set was dropped from the input dataset and an additional comment from each original dataset (one `racist', one `sexist', and one `neither') was added in as a replacement. This replacement was kept consistent each time, for a total of 243 unique comments classified in the few-shot mixed-category setting.

@@

Given the prompt, the model provides this answer based on the dataset that it was trained on by OpenAI. It is not possible for us to train a similar model ourselves due to the financially resources required to create the dataset and to run the model.

@@

Under the few-shot setting, the model is given three examples. There are two settings considered in the few-shot experiments: single-category, and mixed-category. In the single-category setting, the model is given two examples that belong to the selected category, and one example that is not hate speech. The answer is expected to be binary. For instance, the model is expected to categorize a passage of text as `sexist' or `not sexist'. An example of this is:

@@

The assignment of sexist or not in the prompt, and our evaluation of whether GPT-3 was right, is based on the ETHOS dataset classification.

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as `sexist', `racist', or `neither'. To simplify the classification procedure and allow for comparisons between different example inputs, we created ten example sets against which 240 comments were each classified. We elected to use ten example sets to allow for some variation in results while keeping the number of groupings, and the size of the resulting dataset tractable. The example sets each consist of one sexist, one racist, and one `neither' comment and were created using the comments at indexes one through ten from each dataset. Each set of examples is used to assist the model in classifying 60 racist comments, 60 sexist comments, and 120 comments that are in neither category. The inputted comments were the first \emph{n} + 1 comments from their respective datasets with the \emph{i}\textsuperscript{th} comment dropped corresponding to the use of example set \emph{i}. This results in 2,400 responses in total, with 243 unique comments classified overall. The ten sets of examples are shown in Appendix \ref{appendxa}. An example of this is:

@@

\hypertarget{results}{%
\section{Results}\label{results}}

We assess GPT-3's performance in all settings using accuracy, precision, recall, and F1 score. Accuracy is the proportion of correctly classified comments (hate speech and non-hate speech) out of all comments classified. Precision is the proportion of hate speech comments correctly classified out of all comments classified as hate speech (both correctly and incorrectly). Recall is the proportion of hate speech comments correctly classified out of all hate speech comments in the dataset (both correctly and incorrectly classified). The F1 score is the harmonic mean of precision and recall. In the case of hate speech classification, we see it as better to have a model with high recall, meaning a model that can identify a relatively high proportion of the hate speech text within a dataset. But the F1 score can provide a more well-rounded metric for model performance and comparison.

@@

The overall results of the zero-shot experiments are presented in Table \ref{tab:zeroshot-summary}, and Appendix \ref{appendixbzeroshot} provides additional detail. The model has 34 matches (true positives and negatives) and 26 mismatches (false positives and negatives) in the sexist category, and 41 matches and 19 mismatches in the racist category. The model performs more accurately when identifying racist comments, with an accuracy of 68 per cent, compared with identifying sexist comments, with an accuracy of 57 per cent, but the F1 score for each type of hate speech is roughly the same. The overall ratio of matches and mismatches is 75:45. In other words, the accuracy in identifying hate speech in the zero-shot setting is 63 per cent. The model has an F1 score of 81 per cent in this setting, with a notably high recall of 87 per cent.

@@

\caption{\label{tab:zeroshot-summary}Performance of zero-shot learning in hate speech classification}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Category & Accuracy (\%) & Precision (\%) & Recall (\%) & F1 (\%)\\
\midrule
Racism & 68 & 67 & 73 & 70\\
Sexism & 57 & 54 & 100 & 70\\
Overall & 62 & 58 & 87 & 81\\
\bottomrule
\end{tabular}
\end{table}

@@

The results of the one-shot learning experiments are presented in Table \ref{tab:oneshot-summary}, and Appendix \ref{appendixboneshot} provides additional detail. The model has 37 matches and 23 mismatches in the racist category, and 36 matches and 24 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model performs roughly the same in terms of identifying racist comments and identifying sexist comments. The overall ratio of matches and mismatches is 73:47. In other words, the accuracy of identifying hate speech in the one-shot setting is 61 per cent. The general performance in the one-shot setting is worse than in the zero-shot setting, with lower accuracy, recall, precision, and F1 score overall.

@@

\caption{\label{tab:oneshot-summary}Performance of one-shot learning in hate speech classification}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Category & Accuracy (\%) & Precision (\%) & Recall (\%) & F1 (\%)\\
\midrule
Racism & 62 & 59 & 80 & 68\\
Sexism & 60 & 57 & 77 & 66\\
Overall & 61 & 58 & 78 & 65\\
\bottomrule
\end{tabular}
\end{table}

@@

The results of the single-category, few-shot learning, experiments are presented in Table \ref{tab:fewshotsingle-summary}, and Appendix \ref{appendixbfewshotsingle} provides additional detail. The model has 41 matches and 19 mismatches in the racist category, and 42 matches and 18 mismatches in the sexist category. Like the one-shot setting, the model performs almost equally well in terms of identifying racist comments compared with identifying sexist comments. The general performance in the single-category few-shot learning setting is slightly more accurate than performance in other settings, with an accuracy of 69 per cent compared with 61 per cent in the one-shot setting and 63 per cent in the zero-shot setting. But, in this context we see the lowest F1 score of the settings considered thus far at 61 per cent.

@@

\caption{\label{tab:fewshotsingle-summary}Performance of single-category few-shot learning in hate speech classification}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Category & Accuracy (\%) & Precision (\%) & Recall (\%) & F1 (\%)\\
\midrule
Racism & 68 & 66 & 77 & 71\\
Sexism & 70 & 75 & 60 & 67\\
Overall & 69 & 69 & 68 & 61\\
\bottomrule
\end{tabular}
\end{table}

@@

To reduce the chance of the model generating answers that are out of scope, a brief instruction is added to the prompt, specifying that the answers be: `sexist', `racist', or `neither'. The addition of an instruction successfully restricts the generated answers within the specified terms. The unique generated answers are: `racist', `sexist', and `neither'.
